---
title: "00_scratch_r_py.Rmd"
author: "Sarvam"
date: "4/12/2020"
output: html_document
---

#To change the current working directory to specified path

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = normalizePath("/home/sarvammittal/projects/Life_Expectancy/https:/github.com/datascience-practice-projects/Visualizing-Inequalities-in-Life-Expectancy.git"))  
```

```{python}
import os
os.chdir("/home/sarvammittal/projects/Life_Expectancy/https:/github.com/datascience-practice-projects/Visualizing-Inequalities-in-Life-Expectancy.git")
```
# importing libraries

```{r, echo=FALSE}
# R libraries
library(dplyr)
library(tidyr)
library(ggplot2)
```

```{python}
# python libraries

import numpy as np 
import pandas as pd 
import tensorflow as tf
import matplotlib.pyplot as plt
import seaborn as sns

#Ignore Warnings
import warnings
warnings.filterwarnings('ignore')
sns.set_style('whitegrid')
sns.set_context('paper')
```

# Global defaults
```{r}
options(repr.plot.width = 6, repr.plot.height = 6)
```

```{python}
pd.set_option('display.max_colwidth', 1000, 'display.max_rows', None, 'display.max_columns', None)
```

## Utility functions

```{r}
# TODO refactor this 
print_ln <- function() {
  cat(  "--------------------------------------------------------------------------------\n"  )
}
```

```{python}
plt.style.use('ggplot')
sns.set(style='whitegrid')

def print_ln():
    print('-' * 80, '\n')
```


# Life Expectancy

#Life expectancy is a statistical measure of the average time an organism is expected to live, based on the year of its birth, its current age and other demographic factors including gender.


# Exploratory data analysis
```{python}
#reading the csv
```{r}
life_expectancy <- read.csv("../data/raw/UNdata_Export_20200411_102825637.csv", stringsAsFactors = FALSE)
```

```{python}
df = pd.read_csv('./data/raw/UNdata.csv')
df.head()
```

```{python}
list(df.columns)
#Country or Area: Name of the country
#Subgroup: Gender
#Year: Range of Years of a particular country for which data is recorded
#Source: From where this data is taken 
#Unit: Unit of Year column 
#Value: Life expectancy value 
#Value Footnotes : NaN
#We have to remane Country or Area to Country, Subgroup to gender
#We have to drop Unit and Value Footnotes Column
```

```{python}
#renaming the columns
df = df.rename(columns = {"Country or Area": "Country", 
                                  "Subgroup":"Gender"}) 
df.head()
```

```{python}
#dropping the columns which are not of any use
df=df.drop(['Unit','Value Footnotes','Source'], axis = 1) 
df.head()
```

```{python}
#checking the null values
round(100*(df.isnull().sum()/len(df.index)), 2)
#Year and Value columns has 11 values as null.We can leave them as they are very less in ratio leass than 1%
```

```{python}
#it has 1571 rows and 5 columns
df.shape
```
```{python}
#coulmn wise describe
print(df.Country.describe()); print_ln()
print(df.Gender.describe()); print_ln()
print(df.Year.describe()); print_ln()
print(df.Value.describe());print_ln()

```

#Task according to datacamp

#Task1
# Filter life_expectancy to obtain all records such as Year is equal to 2000-2005 and Subset the dataset to include just three columns: Country.or.Area, Subgroup, and Value.Convert Subgroup into two other columns called Female and Male, reshaping dataset from long to wide.Print the first rows of the resulting dataset.

```{r}
subdata <- life_expectancy %>% 
  filter(Year=="2000-2005") %>%
  select(Country.or.Area, Subgroup, Value) %>%
  spread(Subgroup, Value)
```

```{python}
life_expectancy = df.loc[df['Year'] == '2000-2005']
life_expectancy.head()
```
```{python}
#here we have already rename the columns ie Country or Area to Country, Subgroup to Gender
life_expectancy=life_expectancy.filter(["Country", "Gender", "Value"])
life_expectancy.head()
```

```{python}
life_expectancy = life_expectancy.pivot_table(index=['Country'],columns='Gender',values='Value',fill_value=0)
life_expectancy.head()
```

```{python}
#resetting the index of pivot tabel
life_expectancy = life_expectancy.reset_index()
life_expectancy.head()
```
# Task 2
#Use the ggplot function to initialize a ggplot object. Declare subdata as the input data frame and set the aesthetics to represent Male on the x-axis and Female on the y-axis.
#Add a layer to represent observations with points using geom_point.

```{r}
  subdata %>%
  ggplot(aes(x = Male, y = Female)) +
  geom_point() 
```

```{r}
  subdata %>%
  ggplot(aes(x = Male, y = Female)) +
  scale_x_continuous(limits=c(35,85)) +
  scale_y_continuous(limits=c(35,85)) +
  geom_point() + 
  geom_abline(slope = 1, intercept = 0, linetype = 2) + 
  xlab("Males") +
  ylab("Females") +
  labs(title = "Life Expectancy at Birth by Country", caption = "Source: United Nations Statistics Division")
```

```{r}
top_male <- subdata %>% arrange(Male-Female) %>% head(3)
top_female <- subdata %>% arrange(Female-Male) %>% head(3)
ggplot(subdata, aes(x=Male, y=Female, label= Country.or.Area))+
  geom_point(colour="white", fill="chartreuse3", shape=21, alpha=.55, size=5)+
  geom_abline(intercept = 0, slope = 1, linetype=2)+
  geom_text(data=top_male, size=3)+
  geom_text(data=top_female, size=3)+
  theme_bw()+
  scale_x_continuous(limits=c(35,85))+
  scale_y_continuous(limits=c(35,85))+
  labs(title="Life Expectancy at Birth by Country",
       subtitle="Years. Period: 2000-2005. Average.",
       caption="Source: United Nations Statistics Division",
       x="Males",
       y="Females")
```

```{python}
#Interactive graph using plotly
import plotly.express as px

fig = px.scatter(life_expectancy, x="Male", y="Female",
                 hover_name="Country", log_x=True, size_max=60)
fig.show()
```


```{python}
from plotnine import *
p = ggplot(aes(x='Male', y='Female'), data=life_expectancy)
# Add layers
p + geom_point()
# Add a line 
p + geom_point() + geom_line()
# Add a trend line
p + geom_point() + geom_line() + stat_smooth(color='blue')
```

```{python}
ggplot(life_expectancy, aes(x='Male', y='Female')) + \
    geom_point(colour="black", alpha=.55, size=5) + \
    ggtitle("Life Expectancy at Birth by Country")+ \
    geom_abline(slope=1, intercept=0, color='steelblue')
```
#Task3
#Convert Year into four other columns called Female_2000_2005, Female_1985_1990, Male_2000_2005 and Male_1985_1990, reshaping dataset from long to wide.

```{r}
subdata2 <- life_expectancy %>% 
  filter(Year %in% c("1985-1990", "2000-2005")) %>% 
  mutate(Sub_Year=paste(Subgroup, Year, sep="_")) %>% 
  mutate(Sub_Year=gsub("-", "_", Sub_Year)) %>% 
  # the `~` is negation
  select(-Subgroup, -Year)  %>%
  spread(Sub_Year, Value) %>%
  mutate(diff_Female = Female_2000_2005 - Female_1985_1990, diff_Male = Male_2000_2005 - Male_1985_1990 )
```
  
```{python}
# checking the original df
df.head()
```

```{python}
#filtering out 1985-2000 and 2000-2005 data
subsetDataFrame = df[df['Year'].isin(['2000-2005', '1995-2000']) ]
subsetDataFrame.head()
```

```{python}
#subsettibg the dataframe using pivot tables
life_expectancy2 = (subsetDataFrame.pivot_table(index=['Country'], columns=['Gender', 'Year'],fill_value=0))
life_expectancy2.head()
```

```{python}
#flattening the multindex to flaten
life_expectancy2.columns = ["_".join(v) for v in life_expectancy2.columns.values]
life_expectancy2.head()
```

```{python}
#checking the columns of  flaten index
list(life_expectancy2.columns)
```

```{python}
#Converting the pivot tables to dataframe
life_expectancy2=pd.DataFrame(life_expectancy2)
life_expectancy2.head()
```
```{python}
#reseting dataframe index
life_expectancy2 = life_expectancy2.reset_index()
life_expectancy2.head()
```
```{python}
list(life_expectancy2.columns)
```
```{python}
life_expectancy2.head()

```

```{python}
#creating two new columns 
life_expectancy2['diff_Female']= life_expectancy2['Value_Female_2000-2005']-life_expectancy2['Value_Female_1995-2000']
life_expectancy2['diff_Male']= life_expectancy2['Value_Male_2000-2005']-life_expectancy2['Value_Male_1995-2000']
life_expectancy2.head()
```

#Task 4 
#Use the ggplot function to initialize a ggplot object. Declare subdata as the input data frame and set the aesthetics to represent diffMale on the x-axis and diffFemale on the y-axis.
#Add a layer to represent observations with points using geom_point.

```{r}
ggplot(subdata2, aes(x= diff_Male, y=diff_Female, label=Country.or.Area))+
  geom_point(colour="white", fill="chartreuse3", shape=21, alpha=.55, size=5)+
  geom_abline(intercept = 0, slope = 1, linetype=2)+ 
  geom_hline(yintercept = 0, linetype=2)+
  geom_vline(xintercept = 0, linetype=2)+
  scale_x_continuous(limits=c(-25,25))+
  scale_y_continuous(limits=c(-25,25))+
  theme_bw()+
  labs(title="Life Expectancy at Birth by Country in Years",
       subtitle="Difference between 1985-1990 and 2000-2005. Average.",
       caption="Source: United Nations Statistics Division",
       x="Males",
       y="Females")
```


```{r}
top <- subdata2 %>% arrange(diff_Male+diff_Female) %>% head(3)
bottom <- subdata2 %>% arrange(-(diff_Male+diff_Female)) %>% tail(3)
ggplot(subdata2, aes(x=diff_Male, y=diff_Female, label=Country.or.Area), guide=FALSE)+
  geom_point(colour="white", fill="chartreuse3", shape=21, alpha=.55, size=5)+
  geom_abline(intercept = 0, slope = 1, linetype=2)+
  scale_x_continuous(limits=c(-25,25))+
  scale_y_continuous(limits=c(-25,25))+
  geom_hline(yintercept=0, linetype=2)+
  geom_vline(xintercept=0, linetype=2)+
  labs(title="Life Expectancy at Birth by Country",
       subtitle="Years. Difference between 1985-1990 and 2000-2005. Average.",
       caption="Source: United Nations Statistics Division",
       x="Males",
       y="Females")+
  theme_bw()
```

```{python}
plot = sns.scatterplot(x='diff_Male', y='diff_Female', hue=None, style=None, size=None, data=life_expectancy2, palette=None, hue_order=None, hue_norm=None, sizes=None, size_order=None, size_norm=None, markers=True, style_order=None, x_bins=None, y_bins=None, units=None, estimator=None, ci=95, n_boot=1000, alpha='auto', x_jitter=None, y_jitter=None, legend='brief', ax=None)
plot
```
```{python}
#Interactive graph using plotly
import plotly.express as px

fig = px.scatter(life_expectancy2, x="diff_Male", y="diff_Female",
                 hover_name="Country", log_x=True, size_max=60)
fig.show()
```

```{python}
from plotnine import *
p = ggplot(aes(x='diff_Male', y='diff_Female'), data=life_expectancy2)
# Add layers
p + geom_point()
# Add a line 
p + geom_point() + geom_line()
# Add a trend line
p + geom_point() + geom_line() + stat_smooth(color='blue')

```

```{python}

ggplot(life_expectancy2, aes(x='diff_Male', y='diff_Female')) + \
    geom_point(colour="black", alpha=.55, size=5) + \
    ggtitle("Life Expectancy at Birth by Country between 1985-2000")+ \
    geom_abline(slope=1, intercept=0, color='steelblue')
```







